---
title: "Survival Analysis - An Essay for Compressor Failure Prediction"
author: "Dario H. Romero - Oil & Gas Data Scientist"
date: "April 6, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# What is Survival Analysis

Survival analysis is also called duration analysis, transition analysis, failure time analysis, and time-to-event analysis. For the sake of centering the discussion of this topic I would do a brief review of the theory behind, an exercise for survival analysis in medicine and then use that topic for predicting pump failures in an Oil & Gas setup, in particular Compressor Failure Analysis & Prediction.

# Data Set for Analysis

I will be utilizing a Permutational Algorithm for Survival Analysis packaged under the library [PermAlgo](https://cran.r-project.org/web/packages/PermAlgo/PermAlgo.pdf), to generate synthetic data for this anlysis.

# Rationale of Survival Analysis

In Survival Analysis we are concerned in situations where:

 - Subjects are tracked until an event happens (failure) or we lose them from the sample (censored observations).
 - We are interested in how long they stay in the sample (survival).
 - We are also interested in their risk of failure (hazard rates).


# Survival Estimation of time-to-event data as an approximation to Pump Failure Analysis

Survival estimates are an essential compliment to multivariable regression models for
time-to-event data, both for prediction and illustration of covariate effects. They are
easily obtained under the Cox proportional-hazards model (CoxPH).

One of the strengths of the Cox model is its ability to encompass covariates that change over
time. The practical reason that time-dependent covariates work is based on the underlying way
in which the Cox model works: at each event time the program compares the current covariate
values of the subject who had the event to the current values of all others who were at risk at
that time.

For further in depth analysis please take a look at this [document](http://rstudio-pubs-static.s3.amazonaws.com/5896_8f0fed2ccbbd42489276e554a05af87e.html)

## Mathematical Background

Let $Y_i(t), i = 1, . . . , n$ be the indicator that subject $i$ is at risk and under observation at
time $t$. Let $N_{i}(t)$ be the step function for the $i_{th}$ subject, which counts the number of
“events” for that subject up to time $t$. The total number of deaths that have occurred
up to time $t$ will be $\overline{N}(t) = \sum N_i(t)$, and the number of subjects at risk at time $t$ will
be $\overline{Y}(t) = \sum Y_i(t)$. It will also be useful to define $d(t)$ as the number of deaths that
occur exactly at time $t$.

---

## Survival Estimation

The most common estimate of the survival distribution, the `Kaplan-Meier (KM)` estimate,
is a product of survival probabilities:
$$\hat{S}_{KM}(t) =  \prod_{t_i < t}\frac{\overline{Y}(t_i) - d(t_i)}{\overline{Y}(t_i)}$$

Graphically, the `Kaplan-Meier survival curve` appears as a step function with a drop at
each death. Censoring times are often marked on the plot as “+” symbols.


---

## Libraries

```{r libraies}
## Load survival package
library(survival)
## List datasets in survival package
#data(package = "survival")

## Load lung data
data(lung)

## Show first 6 rows
head(lung)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
set.seed(123)
plot(x = lung$time, y = lung$status - 1)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
